### 2016-06-29 doc/publications/plos16/plos16.tex
changes to plos16.tex

While unsupervised clustering is useful for selected clinical applications~\cite{Li2015}, a more common analysis task is
supervised learning, i.e. given a phenotype of clinical label determine the most associated features.  One of the
simplest supervised machine learning methods is logistic regression (LR), which creates a decision boundary by fitting a
linear combination of the feature variables.  Processing the `wide' genomic data poses difficulties for machine learning
applications as having more features $p$ than samples $n$ causes the model to overfit the data and generalize poorly.
This is known as the `curse of dimensionality'~\cite{Bauer2014}.  Random Forests~\cite{Breiman.2001} (RF) as an ensemble of weak learners 
\marginpar{weak learners only in a misuse of the term. ``weak learners'' was applied to decision stumps}
%http://stats.stackexchange.com/questions/77018/random-forest-is-it-a-boosting-algorithm
%http://stats.stackexchange.com/questions/173390/gradient-boosting-tree-vs-random-forest
overcomes this issue and is hence ideally suited to this application.  Furthermore, RF are able to capture interactions
between features, which is of importance for modelling complex polygenic diseases.

@@

>>>>

While unsupervised clustering is useful for selected clinical applications~\cite{Li2015}, a more common analysis task is
supervised learning, i.e. given a phenotype of clinical label determine the most associated features.  One of the
simplest supervised machine learning methods is logistic regression (LR), which creates a decision boundary by fitting a
linear combination of the feature variables.  Processing the `wide' genomic data poses difficulties for machine learning
applications as having more features $p$ than samples $n$ causes the model to overfit the data and generalize poorly.
This is known as the `curse of dimensionality'~\cite{Bauer2014}.  Random Forests~\cite{Breiman.2001} (RF) 
apply the technique of bootstrap aggregation or bagging to decision trees. The training set is randomly sampled $B$
times with replacement. A tree is grown on each of these $B$ samples and the results are combined
\begin{equation*}
{\bar {h}}={\frac {1}{B}}\sum _{b=1}^{B} h(x;\theta_b)
\end{equation*}

Let $PE_f$ be the predicton error for the forest. Then
\begin{equation*}
PE_f \leq \hat{\rho} PR_b
\end{equation*}
where $\hat{\rho}$ is the weighted correlation of the residuals $Y-  h(X;\theta)$ and $Y-  h(X;\theta^\prime)$
for independent $\theta,  \theta^\prime $

So what is required for  accurate random forest regression is low correlation between residuals of differing tree members of
the forest, and (ii) low prediction error for the individual trees.

The strategy employed to achieve these ends is as follows:
1. To keep individual error low, grow trees to maximum depth.
To keep residual correlation low randomize via
a) Grow each tree on a bootstrap sample from the training data.
b) Specify the number of covariates. At each node of every tree
select covariates and pick the best split of that node based on these
covariates.\cite{Segal.2004}

\marginpar{weak learners only in a misuse of the term. ``weak learners'' was applied to decision stumps}
%http://stats.stackexchange.com/questions/77018/random-forest-is-it-a-boosting-algorithm
%http://stats.stackexchange.com/questions/173390/gradient-boosting-tree-vs-random-forest
overcomes this issue and is hence ideally suited to this application.  Furthermore, RF are able to capture interactions
between features, which is of importance for modelling complex polygenic diseases.

@@

\begin{table}
\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{|p{4.2cm}|l|l|l|l|l|l|l|}
\hline
\bf{Genome }  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
\hline
 chr 22 - phase1 (2.1GB)  &  RF (R)\footnote{10 nodes, 500 trees, dpas03} 
                                   &  0.456 & 2.89 & 18GB \\ \hline
\end{tabular}
\medskip
\begin{flushleft} 
\caption{some R results on smaller data sets.}
\label{table:model_outputs}
\end{flushleft}
\end{center}
%\end{sidewaystable}
\end{minipage}
\end{table}

% \begin
% \begin{table}[!ht]
% \caption{
% {\bf Performance comparison between the different Machine Learning algorithms.}}
% \begin{tabular}{|p{4.2cm}|l|l|l|l|l|l|l|}
% \hline
% \bf{Genome }  & \bf{method} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory/Node} \\
% \hline
%  chr 22 - phase1 (2.1GB)  &  RF (R) &  0.456 & 2.89 & 18GB \\ \hline
% \end{tabular}
% \begin{flushleft} 
% \end{flushleft}
% \label{table1}
% %\end{adjustwidth}
% \end{table}
